DRAFT: Scalable, Robust Multipath Communication on the Butterfly Network

### 2016-01-04 netft.tex
Add figure to section multipath fault tolerance.
Add figure showing butterfly network connection rule.
Finish routing subsection.

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{graphicx}
\usepackage{breqn}

@@

\title{DRAFT: Robust Communication in Structured Networks}
\author{
    Edward L. Platt\\
    elplatt@umich.edu
    \and
    Daniel M. Romero\\
    drom@umich.edu
}
\date{December 2015}

@@

We achieve robust communication between trusted neighborhoods through
Concurrent Multipath Routing (CMR) \cite{}.
If there are multiple {\em redundant paths} (i.e., paths sharing no nodes or
links) between Alice and Bob's trusted neighborhoods, multiple copies of a
message can be sent along different paths, enabling standard error detection and
error correction techniques to be applied at the destination.
If Alice randomly selects $k$ of $n$ total redundant paths, Eve's best strategy
is to compromise at least one node on as many of the paths as possible.
Because an error can be detected if even a single copy has been tampered with,
Eve must compromise all $k$ paths in order to convince Bob to accept a compromised
message.
If Eve has the resources to compromise $l$ total nodes, the probability that
Bob receives an undetectable error is:
\beq
\label{eq:pe}
p_e &=& \frac{l!(n-k)!}{n!(l-k)!}.
\eeq
Letting $k=\alpha n$ and $l=\beta n$, then applying Stirling's
approximation gives:
\begin{eqnarray}
\nonumber
p_e &\approx&
\frac{\sqrt{\beta(1-\alpha)}}{\sqrt{\beta-\alpha}} \\
& &
\times {}
\label{eq:pe_approx}
\left[
    \left( \frac{\beta-\alpha}{1-\alpha} \right)^{\alpha}
    \left( \frac{\beta}{\beta-\alpha} \right)^{\beta}
    (1-\alpha)
\right]^n
\end{eqnarray}
Note that Eq. (\ref{eq:pe_approx}) is exponential in the number of redundant
paths $n$, which depends only on the network structure, and not on the
resources available to individual nodes.
If a network has a large number of redundant paths between trusted neighborhoods,
a very low probability of undetected errors can be achieved, even if an adversary
has the resources to compromise a very large fraction of the avialable paths.

@@

>>>>

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{breqn}

@@

\DeclareMathOperator{\wbf}{W-BF}

\title{DRAFT: Scalable, Robust Multipath Communication on the Butterfly Network}
\author{
    Edward L. Platt\\
    elplatt@umich.edu
    \and
    Daniel M. Romero\\
    drom@umich.edu
}
\date{January 2016}

@@

We achieve robust communication between trusted neighborhoods through
Concurrent Multipath Routing (CMR) \cite{}.
If there are multiple {\em redundant paths} (i.e., paths sharing no nodes or
links) between Alice and Bob's trusted neighborhoods, multiple copies of a
message can be sent along different paths, enabling standard error detection and
error correction techniques to be applied at the destination.
If Alice randomly selects $k$ of $n$ total redundant paths, Eve's best strategy
is to compromise at least one node on as many of the paths as possible.
Because an error can be detected if even a single copy has been tampered with,
Eve must compromise all $k$ paths in order to convince Bob to accept a compromised
message.
If Eve has the resources to compromise $l$ total nodes, the probability that
Bob receives an undetectable error is:
\beq
\label{eq:pe}
p_e &=& \frac{l!(n-k)!}{n!(l-k)!}.
\eeq
Letting $k=\alpha n$ and $l=\beta n$, then applying Stirling's
approximation gives:
\begin{eqnarray}
\nonumber
p_e &\approx&
\frac{\sqrt{\beta(1-\alpha)}}{\sqrt{\beta-\alpha}} \\
& &
\times {}
\label{eq:pe_approx}
\left[
    \left( \frac{\beta-\alpha}{1-\alpha} \right)^{\alpha}
    \left( \frac{\beta}{\beta-\alpha} \right)^{\beta}
    (1-\alpha)
\right]^n
\end{eqnarray}
Note that Eq. (\ref{eq:pe_approx}) is exponential in the number of redundant
paths $n$, which depends only on the network structure, and not on the
resources available to individual nodes.
If a network has a large number of redundant paths between trusted neighborhoods,
a very low probability of undetected errors can be achieved, even if an adversary
has the resources to compromise a very large fraction of the avialable paths.
The value of $p_e$ for various $k$, $l$, and $n$ is shown in Figure
\ref{fig:perror}.

\begin{figure}
\includegraphics{fig-perror.pdf}
\caption{
Probability $p_e$ of an undetected error as a function of the number of message
copies $k$ and compromised paths $l$.
\label{fig:perror}
}
\end{figure}

@@

We now describe a Concurrent Multipath Routing scheme on a specific family
of networks: the butterfly network \cite{}.
Several variations on the butterfly network exist.
Specifically, we utilize the wrap-around butterfly.
We define the $m$-dimensional directed wrap-around butterfly $\wbf(m)$:
\beq
\wbf(m) &=& (V, E_\downarrow \cup E_\rightarrow) \\
V &=& \mathbb{Z}_m \times \mathbb{Z}_{2^m} \\
E_\downarrow &=& \{((l,z),(l+1 (\text{mod } m),z) \nonumber \\
&& \; | \, l \in \mathbb{Z}_d, z \in \mathbb{Z}_{2^m}\} \\
E_\rightarrow &=& \{(l,z),(l+1 (\text{mod } m), z \oplus 2^l) \nonumber \\
&& \; | \, l \in \mathbb{Z}_d, z \in \mathbb{Z}_{2^m}\},
\eeq
where $\oplus$ represents the bitwise XOR operator.
Each node is associated with a level $l$ and an $m$-bit integer $z$.
There are two types of edges, shown in Figure \ref{fig:butterfly}.
Down edges ($E_\downarrow$) connect nodes sharing the same $z$ value
in a cycle of increasing level $l$.
Down-right edges ($E_\rightarrow$) also link to a node of level $l + 1$,
but one having the bitstring equal to $z$ with the $l$th bit flipped.

The wrap-around butterfly network has a number of properties making it useful for multipath routing:
\begin{description}
\item[Vertex transitivity]
The problem
of finding a route between arbitrary nodes $v$ and $w$
can be reduced to finding a route from node $(0,0)$ to some $\tilde{w}$.
\item[Logarithmic diameter]
For any two nodes, the length of the shortest path between them is logorithmic
in the size of the network.
\item[Constant degree]
In practical applications, each communication link requires additional resources,
such as physical infrastructure or entries in a routing table.
A constant degree limits the number of such resources needed as the network
grows in size.
\end{description}

\begin{figure}
\begin{center}
\includegraphics{fig-butterfly.pdf}
\end{center}
\caption{
Schematic illustration of the two types of edges in a directed butterfly network.
The node $(l,z)$ is shown as the bit string $z$ with a square around the $l$th bit.
\label{fig:butterfly}
}
\end{figure}

\subsection{Multipath Routing on the Butterfly Network}

We now describe a routing scheme that provides $2^h$ redundant paths between
the $h$-hop trusted neighborhoods of any two nodes in an $m$-bit
wrap-around butterfly network.
Utilizing vertex transitivity, we label the source node as $(0, 0)$ and
denote the destination node as $w = (l_w, z_w)$.

Let $s$ be an integer such that $0 \leq s < 2^h$.
Let $v_s^{(t)} = (l^{(t)}, z^{(t)})$ be the $t$th node in the path labeled by $s$.
For convenience, we will omit the subscript $s$.
We define two partitionings of the integers in $\mathbb{Z}_{2^m}$:
one having the lowest $h$ bits matching the
bits of $s$, and one having the $h$ bits preceeding the destination level $l_w$
matching $s$:
\beq
S_s &=& \{z \in \mathbb{Z}_{2^m} | \forall i \in \mathbb{Z}_h z_i = s_i \} \\
R_s &=& \{z \in \mathbb{Z}_{2^m} | \forall i \in \mathbb{Z}_h z_{(l_w - h + i)} = s_i \}.
\eeq
Note that if $r \neq s$, then $S_s \cap S_r = R_s \cap R_r = \emptyset$.

We now construct a path such that between trusted neighborhoods $z^{(t)}$ is always
in $S_s$, $R_s$, or both, guaranteeing that the path does not overlap with the
other paths $v_r^{(t)}$.
Routing proceeds in stages, with the level $l$ increasing by 1 at each hop.
In Stage 1 ($0 \leq t < h$), down or down-right edges
are chosen such that the the $t$th bit of $z^{(t+1)}$ is equal to the $t$th bit
of $s$. Throughout Stage 1, all nodes are within the sender's trusted neighborhood.
At the end of Stage 1, $z^{(h)} \in S_s$, and $z^{(t)}$ will remain so until the level loops
back to $0$ at $t = m$.

In Stage 2 ($h \leq t < l_w - h$), edges are chosen to make the $t$th bit of
$z^{(t+1)}$ match the $t$th bit of $z_w$.

In Stage 3 ($l_w - h \leq t < l_w$), the bits of $z^{(t)}$ are chosen to match $s$,
such that after the stage is complete, $z^{(l_w)} \in R_s$.

In Stage 4 ($l_w \leq t < m$), as in stage 2,
paths are chosen such that the $t$th bit of $z^{(t+1)}$ matches $z_w$.
After Stage 4, all bits of $z^{(m)}$ are equal to those of $z_w$
except for the first $h$ and the $h$ preceeding index $l_w$.
$z^{(m)}$ is also in both $S_s$ and $R_s$.

At this point, we define $\tau = t - m$.
In Stage 5 ($0 \leq \tau < h$), the first $h$ bits of $z^{(t)}$ are set to
match $z_w$, potentially removing $z^{(t)}$ from $S_s$.

In Stage 6 ($h \leq \tau < l_w - h$), all down edges are chosen, incrementing
the level without any effect on $z^{(t)}$.
At the end of Stage 6, $z^{(m + l_w - h)}$ is still in $R_s$ and
$v^{(m + l_w - h)}$ is now within the trusted neighborhood of $w$.

In the seventh, and final stage ($l_w - h \leq \tau < l_w$), the $h$ bits of $z^{(t)}$
preceeding index $l_w$ are set to match $z_w$.
After this stage, $v^{(m + l_w)} = w$ and routing is complete.
